\section{Introduction}
\label{sec:bg}

Our entire research project has its roots in the study of the dynamics
of morphisms over projective spaces, specifically the periodic points
(or more generally, preperiodic points) of such morphisms. One
profound conjecture in this field is the Uniform Boundedness
Conjecture, put forth by Morton and Silverman in
1994~\cite{MR1264933}, which bounds the number of preperiodic points
of a morphism in terms of the the degree of the base field (over the
rational field $\Q$), the dimension of the projective space, and the
degree of the morphism. The conjecture is strong in the sense that the
dependence on the morphism is the bare minimum --- the coefficients of
the morphism are not referenced at all.

We will define the necessary objects leading up to the Uniform
Boundedness Conjecture, briefly describe the status of the conjecture,
and discuss the specific cases that we will concentrate on. At the end
of this section, our major results will be presented, along with a few
open conjectures.

\begin{definition}
  The \emph{projective space} $\P^N(K)$ of dimension $N$ over a number
  field $K$ is the space of equivalence classes $(K^{N+1} \setminus
  \set{0})/\sim$ in $K^{N+1} \setminus \set{0}$, where the equivalence
  relation $\sim$ is given by
  \[
  (x_0, \dots, x_N) \sim (y_0, \dots, y_N)
  \]
  if and only if $x_i y_j = x_j y_i$ for all $0 \le i, j \le N$. We
  denote such an equivalence class by $[x_0, \dots, x_N]$, where
  $(x_0, \dots, x_N)$ is any element of the equivalence class.
\end{definition}

\begin{definition}
  Given a number field $K$, a \emph{rational map} of degree $d$
  between projective spaces $\P^N(\ol{K})$ and $\P^M(\ol{K})$ is a
  map
  \[
  \begin{gathered}
    \phi: \P^N(\ol{K}) \to \P^M(\ol{K})\\
    P \mapsto [f_0(p), \dots, f_M(p)],
  \end{gathered}
  \]
  where $p \in P$ is any element of the equivalence class $P$, and
  $f_0, \dots, f_M \in \ol{K}[X_0, \dots, X_N]$ are homogeneous
  polynomials of degree $d$ with no common factors. (Well-definedness
  is guaranteed by homogeneity.)

  If all the coefficients of $f_0, \dots, f_M$ lie in $K$, we say that $\phi$ is \emph{defined over} $K$.

  The rational map $\phi$ is \emph{defined at} $P$ if at least one of
  the values $f_0(p), \dots, f_M(p)$ is nonzero.

  The rational map $\phi$ is called a \emph{morphism} if it is defined
  everywhere in $\P^N(\ol{K})$, or equivalently, if the only solution
  in $\ol{K}$ to the simultaneous equations
  \[
  f_0(X_0, \dots, X_N) = \cdots = f_M(X_0, \dots, X_N) = 0
  \]
  is the trivial solution $X_0 = \dots, X_N = 0$.
\end{definition}

\begin{definition}
  Given any morphism $\phi: \P^N \to \P^N$, the \emph{preperiodic
    points of $\phi$} in over some field $K$ is the collection of
  points in $\P^N(K)$ with finite forward orbits, or equivalently,
  \[
  \preper(\phi, \P^N(K)) = \set{P \in \P^N(K) : \text{exist $n_1, n_2
      \in \N$ such that $\phi^{n_1}(P) = \phi^{n_2}(P)$}}.
  \]
\end{definition}

Now we are ready to state the Uniform Boundedness Conjecture.

\begin{conjecture}[Morton-Silverman, 1994 ~\cite{MR1264933}]
  Fix integers $d \ge 2$, $N \ge 1$, and $D \ge 1$. There is a
  constant $C(d, N, D)$ such that for all number fields $K/\Q$ of
  degree at most $D$ and all morphisms $\P^N \to \P^N$ of degree $d$
  defined over $K$,
  \[
  \#\preper(\phi, \P^N(K)) \le C(d, N, D).
  \]
\end{conjecture}

Very little is known about this conjecture. In fact, even the simplest
case $(d, N, D) = (2, 1, 1)$ is not known, that is, the problem of
bounding the number of rational preperiodic points of degree-2
rational morphisms. If we specialize to degree-2 polynomials (i.e.,
maps of the type $[X_0, X_1] \mapsto [a X_0^2 + b X_0 X_1 + c X_1^2,
X_1^2]$, conveniently written as $a z^2 + b z + c$), note that $a z^2
+ b z + c$ can be reduced to the form $z^2 + c'$ by linear
conjugation, which preserves the sizes of orbits, so it suffices to
consider degree-2 polynomials of the type $\phi_c(z) = z^2 + c$. In
this case of $\phi_c(z) = z^2 + c$, the uniform bound is again not
known, but there are partial results for specific small periods.

One can easily show that there are one-parameter families of
$c$-values for which $\phi_c(z)$ has rational periodic point(s) of
exact period 1, 2, or 3. It was shown by Morton~\cite{MR1665198} that
$\phi_c$ cannot have rational periodic points of exact period 4. The
non-existence result was extended to exact period 5 by Flynn, Poonen,
and Schaefer~\cite{MR1480542}, and to exact period 6 (conditional on
the Birch and Swinnerton-Dyer conjecture for a specific abelian
variety) by Stoll~\cite{MR2465796}.

We are interested in generalizing the aforementioned results for
quadratic polynomials and specific small periods to quadratic number
fields, i.e., the $D = 2$ case. For a more detailed discussion of
other takes on the Uniform Boundedness Conjecture, along with a list
of references, see Section 3.3 of \cite{MR2316407}.

From now on, unless otherwise stated, the dimension of the projective
space will always be 1, and we reserve the letter $N$ for the exact
period of periodic points.

\subsection{Main results}
\label{subsec:results}

Below are our main results.

For $N = 4$, by putting together and re-interpreting others' results,
we have

\begin{theorem}
  \label{th:n=4-infinite}
  There are infinitely many rational values $c$ such that $\phi_c(z) =
  z^2 + c$ has periodic points of exact period 4 defined in some
  quadratic extension to $\Q$. All such $c$ are given by
  \[
  c = \frac{1 - 4t^3 - t^6}{4t^2(t^2 - 1)}
  \]
  with $t \in \Q$.

  Moreover, for each such $c \in \Q$, $\phi_c(z)$ has exactly four
  such periodic points.
\end{theorem}

For $N = 5$, we have

\begin{theorem}
  \label{th:n=5-finite}
  There are finitely many rational values $c$ such that $\phi_c(z) =
  z^2 + c$ has periodic points of exact period 5 defined in some
  quadratic extension to $\Q$.
\end{theorem}

In fact, all empirical evidence points to the following stronger
conjecture:

\begin{conjecture}
  \label{cj:n=5-zero}
  There are no rational values $c$ such that $\phi_c(z) = z^2 + c$ has
  periodic points of exact period 5 defined in some quadratic
  extension to $\Q$.
\end{conjecture}

One general result we produced is that

\newcommand{\nd}{\frac{N}{(N, d)}}

\begin{theorem}
  Let $N \in \N^*$, $c \in \Q$, $K$ be a Galois extension of $\Q$ with
  degree $d = [K : \Q]$, and $\set{z_0, \dots, z_{N-1}} \subset K$ be
  an exact $N$-cycle of $\phi_c(z) = z^2 + c$. Then exactly one of the
  following holds:
  \begin{enumerate}[(i)]
  \item $z_{i\nd} = \t(z_0)$ for some $0 \le i \le (N, d)-1$ and some
    nontrivial $\t \in \gal(K/\Q)$;

  \item $\set{z_0, \dots, z_{N-1}} \cap \set{\t(z_0), \dots,
      \t(z_{N-1})} = \es$ for all nontrivial $\t \in \gal(K/\Q)$.
  \end{enumerate}
\end{theorem}

In fact, we have good reasons to believe that the second case never
occurs, hence the conjecture

\begin{conjecture}
  \label{cj:galois-conjugate}
  Let $N \in \N^*$, $c \in \Q$, $K$ be a Galois extension of $\Q$ with
  degree $d = [K : \Q]$, and $\set{z_0, \dots, z_{N-1}} \subset K$ be
  an exact $N$-cycle of $\phi_c(z) = z^2 + c$. Then $z_{i\nd} =
  \t(z_0)$ for some $0 \le i \le (N, d)-1$ and some nontrivial $\t \in
  \gal(K/\Q)$.
\end{conjecture}

Assuming this conjecture, some problems of finding quadratic points
(or even higher) may be reduced to finding rational points. In
particular, it turns out that this conjecture implies
Conjecture~\ref{cj:n=5-zero}.

Before we close the introduction section, note that we have a handful
of programs, written in C++, Mathematica, or Sage, that accompany this
report. These programs are available as a repository on GitHub; see
\cite{src} in the reference list. Whenever we need to refer to such a
program, we print its path in the repository in monospaced font; for
instance, \texttt{mma/abc.nb} refers to the Mathematica notebook
\texttt{abc.nb} in the \texttt{mma} directory of the repository.

%%% Local Variables:
%%% TeX-master: "report"
%%% End:
