\section{Background}
\label{sec:pre}

In this section, we introduce some preliminary notions and tools
necessary for further development of the subject. Readers already
familiar with the concepts introduced may safely skip this section.

\subsection{Genus of curves}

First we sketch the notions of the normalization and genus of a
projective curve.

Normalization is a widely used technique in algebraic geometry to
desingularize algebraic curves (and more generally, schemes). For each
curve $C$ with finitely many singularities, there is a canonical way
to construct a nonsingular curve $\tilde{C}$, called the
\emph{normalization of $C$}, along with a surjection $\varphi:
\tilde{C} \to C$, called the \emph{normalization morphism}, such that
$\varphi$ is an isomorphism outside the finitely many singularities of
$C$. For a rigorous treatment of normalization, see Subsection 4.1.2
of \cite{MR1917232}.

There are two different notions of genus, the \emph{arithmetic genus}
and the \emph{geometric genus}.

\begin{definition}
  Let $X$ be a projective curve over a field $K$. The \emph{arithmetic
    genus} of $X$ is defined to be the integer
  \[
  p_a(X) = 1 - \chi_K(\O_X),
  \]
  where $\chi_K$ denotes the Euler-Poincar\'e characteristic of a
  coherent sheaf (see p.~205 of \cite{MR1917232}), and $\O_X$ denotes
  the structure sheaf (see p.~37 of \cite{MR1917232}).
\end{definition}

\begin{definition}
  Let $Y$ be a smooth projective curve over a field $K$. The
  \emph{geometric genus} of $Y$ is defined to be
  \[
  p_g(Y) = \dim_K H^0(Y, \omega_{Y/K}),
  \]
  where $H^0$ denotes the zeroth cohomology group (see p.~182 of
  \cite{MR1917232}), and $\omega_{Y/K}$ denotes the canonical sheaf
  (see p.~239 of \cite{MR1917232}).
\end{definition}

\begin{remark}
  \label{rem:abuse}
  While the arithmetic genus is defined for all projective curves, in
  the above definition the geometric genus is only defined for smooth
  curves. Nevertheless, by common abuse of the concept, we define the
  \emph{geometric genus of any curve $X$} to be the geometric genus of
  its normalization (which is smooth).
\end{remark}

It turns out that for $X$ smooth and geometrically connected, we have
$p_a(X) = p_g(X)$ (see Subsection 7.3.2 of \cite{MR1917232}). In
particular, for the normalization of a curve $C$, the arithmetic genus
and the geometric genus coincide. Therefore, from now on we use the
following notation:

\begin{definition}
  Let $C$ be a projective curve over a field $K$ (not necessarily
  smooth). Then we define $g(C)$ to be the \emph{geometric genus} of a
  curve $C$ (which, recall from Remark~\ref{rem:abuse}, is actually
  the geometric genus of the normalization of $C$.)
\end{definition}

\begin{remark}
  It is worth noting that $g(C)$ can be computed reasonably
  efficiently, provided that we have an explicit equation for $C$. For
  instance, if the polynomial equation $p(x, y) = 0$ defines an affine
  model of $C$ (not necessarily smooth), then we may feed the
  polynomial $p(x, y)$ into some computer algebra system to obtain
  $g(C)$. Computer algebra systems capable of computing the geometric
  genus of an algebraic curve include, but are not limited to, Magma
  and Sage (using its interface to Singular).\footnote{%
    These computer algebra systems are also capable of computing the
    arithmetic genus of algebraic curves, but that is not very
    relevant to our usage cases, as we are mostly interested in the
    normalization (see Theorem~\ref{th:faltings} and related
    discussions).}
\end{remark}

The reason that we are interested in the geometric genus of a curve is
due to the following theorem.

\begin{theorem}[Faltings, 1983 \cite{MR718935}]
  \label{th:faltings}
  Let $C$ be a non-singular algebraic curve of genus $g$ over
  $\Q$. Then the set of rational points $C(\Q)$ on $C$ satisfy:
  \begin{enumerate}
  \item If $g = 0$, then either $C(\Q) = \es$, or $|C(\Q)| = \infty$,
    in which case $C$ is isomorphic over $\Q$ to the projective line
    $\P^1$;

  \item If $g = 1$, then either $C(\Q) = \es$, or $C$ is an
    \emph{elliptic curve}, in which case $C(\Q)$ is a finitely
    generated abelian group;

  \item If $g \ge 2$, then $|C(\Q)| < \infty$.
  \end{enumerate}
\end{theorem}

Faltings' theorem shows that the genus of a curve gives us information
about the number of rational points on it. In particular, there are
only finitely many rational points on a high genus curve (from now on
we refer to a curve with geometric genus $g \ge 2$ as a \emph{high
  genus curve}). Note that this works for any high genus curve
(singular or not), since once we have the finitely many rational
points on the normalization, it is fairly easy to investigate what
happens at each of the finitely many singularities.

\begin{remark}
  Faltings' theorem is ineffective, since it does not provide an
  explicit bound on the number of rational points; nor does it present
  a method to find such a bound.

  Under certain circumstances, there are methods to find an explicit
  bound (usually not sharp), most notably Chabauty and Coleman's
  method \cite{MR808103}.
\end{remark}

\subsection{Resultant of polynomials}

Another tool that we need is the resultant of two polynomials.

\begin{proposition}
  \label{prop:res}
  Let
  \[
  \begin{gathered}
    A(X) = a_0 X^n + a_1 X^{n-1} + \cdots + a_n,\\
    B(X) = b_0 X^m + b_1 X^{m-1} + \cdots + b_m
  \end{gathered}
  \]
  be polynomials of degrees $n$ and $m$ with coefficients
  in a field $k$. There exists a polynomial
  \[
  \res(a_0, \dots, a_n, b_0, \dots, b_m) \in \Z[a_0, \dots, a_n, b_0,
  \dots, b_m],
  \]
  in the coefficients of $A$ and $B$, called the \emph{resultant of
    $A$ and $B$}, with the following properties:
  \begin{enumerate}[(i)]
  \item $\res(A, B) = 0$ if and only if $A$ and $B$ have a common root
    in $\ol{k}$;

  \item If $a_0 b_0 \ne 0$ and if we factor $A$ and $B$ as
    \[
    A = a_0 \prod_{i=1}^n (X - \alpha_i),\,
    B = b_0 \prod_{i=1}^m (X - \beta_j),
    \]
  \end{enumerate}
  then
  \[
  \res(A, B) = a_0^m b_0^n \prod_{i=1}^n \prod_{j=1}^m (\alpha_i -
  \beta_j).
  \]
\end{proposition}

For a proof of this proposition, see Section~2.4 of \cite{MR2316407}.

\begin{remark}
  Since $\res(A, B)$ is a polynomial in the coefficients of $A$ and
  $B$, when the dimensions $n$ and $m$ are reasonably small, $\res(A,
  B)$ can be computed efficiently by computer algebra
  systems. Computer algebra systems capable of computing the resultant
  of two polynomials include, but are not limited to, Magma,
  Mathematica, and Sage.

  As the resultant can be computed efficiently, it is a good way to
  reduce the number of variables in a set of simultaneous polynomial
  equations. For instance, if we have two polynomial equations $A(X,
  Y) = 0$ and $B(X, Y) = 0$, then we may take the resultant with
  respect to $X$ (viewing $X$ as the variable in the above
  definition). Denote it by $\res_X(A, B)$ (we will continue to use
  the $\res_X$ notation when there are more than one variables). Then
  by Property~(i) of the resultant, there is an $X$ solving both $A$
  and $B$ if and only if $Y$ solves $\res_X(A, B)$, which is a
  polynomial in $Y$, since $A$ and $B$ both have coefficients in
  $k[Y]$ when viewed as polynomials in $X$. Therefore, we reduced two
  equations in two unknowns to one equation in one unknown. Similar
  argument applies when there are more variables.
\end{remark}

%%% Local Variables:
%%% TeX-master: "report"
%%% End:
